---
title: ""
format: dashboard
---

# Syllabus and Course Links

```{r include=F}
#| label: setup

options(scipen=100)                         # suppress scientific notation
                                            # install pacman if needed
if (!require("pacman")) install.packages("pacman", repos = "http://lib.stat.cmu.edu/R/CRAN/")
pacman::p_load(pacman, tidyverse, ggthemes, gridExtra, magrittr,
               kableExtra, gt)

#pacman::p_loaded()

```

```{r echo=F}
#| label: created formatted date
# create variables for inline code
today <- Sys.Date()                                                # saves today's date
today_fmt <- stamp("Saturday, January 1, 1999", quiet = T)(today)  # formats date

```

## Row {height="30%"}

### Column

**Professor Pooler (pspooler\@syr.edu) - Office: WSOM Room
518**

-   In-Person Hours: Tue. Thu. 5:00 PM - 5:45 PM
-   Zoom Hours: Mon. Wed. 4:30 PM - 5:30 PM - [**Zoom Hrs.
    Link**](https://syracuseuniversity.zoom.us/j/96889308509){target="_blank"}
-   Also available by appointment on Zoom or in-person

### Column

**Teaching Assistant: Available by Appt.**

-   Arjun Pendharkar - apendhar\@syr.edu

## Row {height="70%"}

[**Spring 2026
Syllabus**](https://docs.google.com/document/d/1bqsRpi8fy_3lC846C9mSVjFgAE9srsrssxSqSbIqWw8/edit?usp=sharing){target="_blank"}

-   NOTE: The order of the material varies slightly between
    semesters so the subjects covered in the syllabus may be
    updated as the semester progresses. I will always let
    you know if there are substantial changes.

[**Spring 2026 Absence Explanation
Form**](https://forms.gle/kuZi7aoLyEcBhqfK8){target="_blank"}

-   If you are absent or have a temporary technical issue
    with POLL EVERYWHERE, use this Google form to explain.

[**Spring 2026 BUA 455 Blackboard
Site**](https://blackboard.syracuse.edu/ultra/courses/_555184_1/outline){target="_blank"}

-   Engagement Questions, Assignments, and Tests will be
    submitted on Blackboard.

[**White Paper about Analytics
Skillset**](https://drive.google.com/file/d/1Beg58FM6cPvPc_8s5eHXWELkXcE3qiVj/view?usp=sharing){target="_blank"}

-   Starting on page 8 (10th document page), this paper
    provides specific language about different roles in data
    analytics.

-   This section may be helpful as you try to explain your
    skillset in interviews.

[**Read this letter about
AI**](https://www.nytimes.com/2025/06/18/learning/we-need-to-chat-gpt.html?smid=url-share){target="_blank"}

-   Using AI in coding and writing is commonplace and
    helpful if used cautiously and ethically.

-   This is a great 'letter' from a high school student who
    'broke up' with AI because she felt it was supplanting
    her own voice. It's worth your time to read and consider
    it.

[**Permananent BUA 455
Website**](https://peneloopy.github.io/bua_455_perm/){target="_blank"}

-   This link provides permanent access to the most recent
    lecture slides, notes, and R projects after the semester
    has ended.

-   Homework assignments, practice questions, and project
    information are not included on the permanent site.

**Required Software Links**

-   [R](https://cran.r-project.org/){target="_blank"} -
    Required all semester

-   [RStudio](https://posit.co/download/rstudio-desktop/){target="_blank"} -
    Required all semester

-   [Quarto](https://quarto.org/docs/get-started/){target="_blank"} -
    Required after HW 1

-   [Posit Cloud](https://posit.cloud/){target="_blank"} -
    Required after Fall Break but can be used sooner if
    needed.

    -   [Student Subscription required for project after
        Fall
        Break](https://posit.cloud/plans){target="_blank"}.

<br>

-   [RPubs](https://rpubs.com/){target="_blank"} - Required
    after Quiz 2

**Software Training**

-   [LinkedIn
    Learning](https://linkedinlearning.syr.edu/){target="_blank"} -
    Free for Syracuse University Students, Staff, and
    Faculty

    -   [R for Data Science: Analysis and
        Visualization](https://www.linkedin.com/learning/r-for-data-science-analysis-and-visualization?u=76816226){target="_blank"}

        -   This training is optional and provides a good
            introduction to R and RStudio which are used in
            this course.

    -   [Python Essential
        Training](https://www.linkedin.com/learning/python-essential-training-18764650?u=76816226){target="_blank"}

        -   Python is not used in this course but it is
            helpful for analytics and data science.

<br>

-   [DataCamp](https://www.datacamp.com/){target="_blank"}

    -   An excellent (but not free) online learning center
        for many software packages. Costs can be subsidized
        (see below).

<br>

-   [Whitman WIRE
    Initiative](https://whitman.syracuse.edu/experiential-learning/wire-initiative){target="_blank"}

    -   The WIRE initiative will subsidize costs of online
        software training that results in certification.

    -   Contact their office to arrange a meeting and
        develop a plan.

<br>

-   [SU Open Source Program Office
    (OSPO)](https://opensource.syracuse.edu/){target="_blank"}

    -   OSPO offers workshops in open source software such
        as R/RStudio, Python, Github and facilitates
        communication between open source software users on
        campus.

# Lecture Material

## Row

[***Poll
Everywhere***](https://pollev.com/penelopepoolereisenbies685){target="_blank"
style="color: lightgreen"}

My User Name: **penelopepoolereisenbies685**

<br>

**This website was last updated on `r today_fmt`.**

-   **Week 7: Time Series, Data Formats, Output Formats,
    Project Introduction**

    -   [**Slides**](https://rpubs.com/PeneLope_PE/1353094){target="_blank"}

    -   [**Notes**](https://rpubs.com/PeneLope_PE/1401052){target="_blank"}

    -   [**Notes
        PDF**](https://drive.google.com/file/d/1M0POE7sdSIcB9CT7ZtSAjymCm63f6x-A/view?usp=sharing){target="_blank"}

    -   [**Week 7 Zipped R Project to
        Download**](https://drive.google.com/file/d/1-4eMBxDAzWhsjrL2yXzylF6kfAFOuq4K/view?usp=sharing){target="_blank"}
        
    -   **Video Recordings**

        -   [**Lecture 13 - 2026-02-24 - 12:30
            PM**](https://video.syr.edu/media/t/1_pedfsimr/398925753){target="_blank"}
            
        -   [**Lecture 14 - 2026-02-26 - 12:30
            PM**](https://video.syr.edu/media/t/1_7xlv11n9/398925753){target="_blank"}

<br>

-   **Week 6: Reshaping. Cleaning, and Joining Data**

    -   [**Slides**](https://rpubs.com/PeneLope_PE/1226774){target="_blank"}

    -   [**Notes**](https://rpubs.com/PeneLope_PE/1226777){target="_blank"}

    -   [**Notes
        PDF**](https://drive.google.com/file/d/1LUym1aUPRd3rvv0DMkJ7ym5GAZhDBJKG/view?usp=sharing){target="_blank"}

    -   [**Week 6 Zipped R Project to
        Download**](https://drive.google.com/file/d/1LSv9HWSMFDj6duRf521CvqbOVHW5JnLt/view?usp=sharing){target="_blank"}

    -   **Video Recordings**

        -   [**Lecture 11 - 2026-02-17 - 12:30
            PM**](https://video.syr.edu/media/t/1_gzwsskij/398925753){target="_blank"}

        -   [**Lecture 12 - 2026-02-19 - 12:30
            PM**](https://video.syr.edu/media/t/1_ln84wzga/398925753){target="_blank"}

<br>

-   **Week 5: Cheat Sheets, AI, Functions, and Review**

    -   [**Slides**](https://rpubs.com/PeneLope_PE/1342482){target="_blank"}

    -   [**Notes**](https://rpubs.com/PeneLope_PE/1223849){target="_blank"}

    -   [**Notes
        PDF**](https://drive.google.com/file/d/1JQPZpqVtGhy_Pka3rvbZ1ekmks0r6CPO/view?usp=sharing){target="_blank"}

    -   [**Week 5 Zipped R Project to
        Download**](https://drive.google.com/file/d/16OOunZmUWf-SgBThe89KBmaeEi9j24l4/view?usp=sharing){target="_blank"}

    -   **Video Recordings**

        -   [**Lecture 9 - 2026-02-10 - 12:30
            PM**](https://video.syr.edu/media/t/1_hq6zko87/398925753){target="_blank"}

<br>

-   **Week 4: Combining Skills and Quiz Review**

    -   [**Slides**](https://rpubs.com/PeneLope_PE/1221027){target="_blank"}

    -   [**Notes**](https://rpubs.com/PeneLope_PE/1221030){target="_blank"}

    -   [**Notes
        PDF**](https://drive.google.com/file/d/1Irb5c6TayC-NyqvOF6_PsQIHxwteM01H/view?usp=sharing){target="_blank"}

    -   [**Week 4 Zipped R Project to
        Download**](https://drive.google.com/file/d/1zLMDnxcfStFY8VPmpK57rN7_Uc4YniCp/view?usp=sharing){target="_blank"}

        -   Unzip and save unzipped project folder to BUA
            455 folder on your desktop.

    -   **Video Recordings**

        -   [**Lecture 7 - 2026-02-03 - 12:30
            PM**](https://video.syr.edu/media/t/1_motqh4v7/398925753){target="_blank"}

        -   [**Lecture 8 - 2026-02-05 - 12:30
            PM**](https://video.syr.edu/media/t/1_zrk24x85/398925753){target="_blank"}

<br>

-   **Week 3: Reshaping and Summarizing Data**

    -   [**Slides**](https://rpubs.com/PeneLope_PE/1217848){target="_blank"}

    -   [**Notes**](https://rpubs.com/PeneLope_PE/1217850){target="_blank"}

    -   [**Notes
        PDF**](https://drive.google.com/file/d/1IpThEkRuYpME6PIdv_j9pOQ5gUpfaeuj/view?usp=sharing){target="_blank"}

    -   [**Week 3 Zipped R Project to
        Download**](https://drive.google.com/file/d/1oLSUXtXYeHnsUdrkn3SpYZDU0NZdZNTc/view?usp=sharing){target="_blank"}

        -   Unzip and save unzipped project folder to BUA
            455 folder on your desktop.

    -   **Video Recordings**

        -   [**Lecture 5 - 2026-01-27 - 12:30
            PM**](https://video.syr.edu/media/t/1_axur1z2s/398925753){target="_blank"}

        -   [**Lecture 6 - 2026-01-29 - 12:30
            PM**](https://video.syr.edu/media/t/1_ruji5fox/398925753){target="_blank"}

<br>

-   **Week 2: Introduction to dplyr commands and ggplot**

    -   [**Slides**](https://rpubs.com/PeneLope_PE/1215286){target="_blank"}

    -   [**Notes**](https://rpubs.com/PeneLope_PE/1215287){target="_blank"}

    -   [**Notes
        PDF**](https://drive.google.com/file/d/1IngBTEB-GuZJawBSBtD6ICc63CZRylEH/view?usp=sharing){target="_blank"}

    -   [**Week 2 Zipped R Project to
        Download**](https://drive.google.com/file/d/1o3GBPckHWmb3dCpk91f-9vqlRuXdwWet/view?usp=sharing){target="_blank"}

        -   Unzip and save unzipped project folder to BUA
            455 folder on your desktop.

    -   **Video Recordings**

        -   [**Lecture 3 - 2026-01-20 - 12:30
            PM**](https://video.syr.edu/media/t/1_6vto5gkp/398925753){target="_blank"}

        -   [**Lecture 4 - 2026-01-22 - 12:30
            PM**](https://video.syr.edu/media/t/1_qm9t4wmo/398925753){target="_blank"}

<br>

-   **Week 1: R Projects, R Quarto Files, and R Syntax**

    -   [**Slides**](https://rpubs.com/PeneLope_PE/1188589){target="_blank"}

    -   [**Notes**](https://rpubs.com/PeneLope_PE/1212697){target="_blank"}

    -   [**Notes
        PDF**](https://drive.google.com/file/d/1ItvsDJaWbEDFNiFbxCcbwQvlgfraRCJR/view?usp=sharing){target="_blank"}

    -   [**Week 1 Zipped R Project to
        Download**](https://drive.google.com/file/d/1nrBPZEMahZ08pAuLYcvLtvYPiAkn2R7D/view?usp=sharing){target="_blank"}

        -   Unzip and save unzipped project folder to BUA
            455 folder on your desktop.

    -   **Video Recordings**

        -   [**Lecture 1 - 2026-01-13 - 12:30
            PM**](https://video.syr.edu/media/t/1_jrqevb76/398925753){target="_blank"}

        -   [**Lecture 2 - 2026-01-15 - 12:30
            PM**](https://video.syr.edu/media/t/1_u0wrjvol/398925753){target="_blank"}

        -   **NOTE:** In response to a request, captions
            will be included in videos going forward. Please
            note that the captions are 96-97% accurate, but
            not perfect and the accuracy may decrease when I
            use technical terminology. Please let me know if
            you have a question.

# Homework Assignments

## Row

-   **Homework 4:**

    -   [**HW 4
        Instructions**](https://rpubs.com/PeneLope_PE/1226728){target="_blank"}

    -   [**HW4_Template.qmd**](https://drive.google.com/file/d/1DFE4VnAdDAmhqIKuSJictN7Ilk6bIZRw/view?usp=sharing){target="_blank"}

    -   [**DATA: HW4_data_files.zip - Unzip and save 3 data
        files to `data`
        folder**](https://drive.google.com/file/d/16lmXwBw3H9_9kiYjXzTupJOrMIN84eh1/view?usp=sharing){target="_blank"}

    -   [**R Project - README
        Template.txt**](https://drive.google.com/file/d/1o0bmzrKnRqAPw6cD8WYvIUGB10Gq71iX/view?usp=sharing){target="_blank"}

    -   [**HW 4 Demo Video
        Playlist**](https://video.syr.edu/playlist/dedicated/353201892/1_yc5vl2le/){target="_blank"}

<br>

-   **Homework 3:**

    -   [**HW 3
        Instructions**](https://rpubs.com/PeneLope_PE/1216489){target="_blank"}

    -   [**HW3_Template.qmd**](https://drive.google.com/file/d/1oFMkCRKE1qWB7GEOIlajK0psDVyAzHs0/view?usp=sharing){target="_blank"}

    -   [**DATA: Box_Office_Mojo_Week3_HW3.csv - Save to
        `data`
        folder**](https://drive.google.com/file/d/1oB_nLuLZ_KagGeCkrsqmGs1tFZGrHw1l/view?usp=sharing){target="_blank"}

    -   [**R Project - README
        Template.txt**](https://drive.google.com/file/d/1o0bmzrKnRqAPw6cD8WYvIUGB10Gq71iX/view?usp=sharing){target="_blank"}

    -   [**HW 3 Demo Video
        Playlist**](https://video.syr.edu/playlist/dedicated/353201892/1_b5xeucc6/){target="_blank"}

<br>

-   **Homework 2:**

    -   [**HW 2
        Instructions**](https://rpubs.com/PeneLope_PE/1214518){target="_blank"}

    -   [**HW2_Template.qmd**](https://drive.google.com/file/d/1o-kM0KC3qiEOwHWmw7I7GL8daGvOAGjd/view?usp=sharing){target="_blank"}

    -   [**R Project - README
        Template.txt**](https://drive.google.com/file/d/1o0bmzrKnRqAPw6cD8WYvIUGB10Gq71iX/view?usp=sharing){target="_blank"}

    -   [**HW 2 Demo Video
        Playlist**](https://video.syr.edu/playlist/dedicated/353201892/1_4ut6eiym/){target="_blank"}

<br>

-   **Homework 1 - Parts 2 and 3:**

    -   [**HW 1
        Instructions**](https://rpubs.com/PeneLope_PE/1212680){target="_blank"}

    -   [**HW1_SETUP.txt**](https://drive.google.com/file/d/1njqHffuwPFfCh1TO-3d_2DRFyPelkGAG/view?usp=sharing){target="_blank"}

    -   [**HW 1 Demo Video
        Playlist**](https://video.syr.edu/playlist/dedicated/353201892/1_epecznax/){target="_blank"}

# Dashboard Project

## Row

[**Full Project Description and Grading
Criteria**](https://docs.google.com/document/d/1LOqbiRK2eOG6F4sKpnTBZ8uti0XqruSBkksp1ChW2UA/edit?usp=sharing){target="_blank"}

## Row

**Purpose(s):**

-   This Data Dashboard group project (23% of course grade)
    serves two purposes in the BUA 455 curriculum:

    1.  Students will better understand the whole data
        management process and how to make good choices at
        each step to make the work process more efficient
        and reproducible.

    2.  Students will learn about the interactive components
        of a data management and visualization project,
        i.e., a dashboard, including how to communicate with
        technically knowledgeable colleagues and
        non-technical audiences.

## Row

**Description of Project:**

-   Student groups will develop a Data Dashboard.

-   In class, we will review how to use the [**Quarto
    Dashboard
    format**](https://quarto.org/docs/gallery/#dashboards){target="_blank"}
    to publish a multi-page dashboard.

-   Each group will aggregate, manage, and curate data for a
    hypothetical client to address their data questions and
    needs.

-   One person in each group should ideally will oversee the
    project as a whole and ensure that the dashboard is
    cohesive.

-   Each page can present a single plot, a set of plots, or
    a plot and a table, and should have explanatory text and
    citations for data sources if needed.

-   The final page of your dashboard will include all
    citations for software, software packages, AI sources,
    and data used to create your dashboard. AI sources, and
    data used to create your dashboard.

# Interesting Data

## Row

**NOTES:**

-   Many data sources provide data in a downloadable format
    such as Excel (.xlsx) or .csv. I recommend saving data
    as CSV UTF-8 files with a .csv postscript.

-   The list below is in alphabetical order and is not an
    exhaustive list.

-   Some data sources do not provide data in a downloadable
    format so the data values have to be scraped from the
    website.

## Row

[**Billboard**](https://www.billboard.com/){target="_blank"}

-   Scraping required

[**BEA - Bureau of Economic
Analysis**](https://www.bea.gov/){target="_blank"}

<br>

[**BLS- Bureau of Labor
Statistics**](https://www.bls.gov/){target="_blank"}

<br>

[**Box Office
Mojo**](https://www.boxofficemojo.com/){target="_blank"}

-   Scraping required but it's straightforward.

[**Competence Centre on Composite Indicators and
Scoreboards**](https://composite-indicators.jrc.ec.europa.eu/){target="_blank"}

-   Great source for unusual datas

[**Data World**](https://data.world/home/){target="_blank"}

-   An ecclectic library of data that can be searched by
    topic of interest

[**Department of Energy - Prices &
Trends**](https://www.energy.gov/energy-economy/prices-trends){target="_blank"}

<br>

[**Eurostat**](https://ec.europa.eu/eurostat){target="_blank"}

-   A wide variety of data about Europe can be accessed at
    this site.

-   Aanalogous to US.Gov for the United States.

[**Federal Reserve
Data**](https://www.federalreserve.gov/data.htm){target="_blank"}

<br>

[**Kaggle**](https://www.kaggle.com/){target="_blank"}

-   An ecclectic library of data that can be searched by
    topic of interest
-   Kaggle also holds competitions where data scientists can
    win money and build a reputation

[**Our World in
Data**](https://ourworldindata.org/){target="_blank"}

-   Good source for geographic and demographic data

[**Sports
Reference**](https://www.sports-reference.com/){target="_blank"}

-   Scraping required
-   Wide range of sports data

[**Statista**](https://www-statista-com.libezproxy2.syr.edu/){target="_blank"}

-   Syracuse University Libraries provides paid membership
    access
-   Use link above and then sign in with SU NetID and
    Password
-   Excellent resource for public and private sector data

[**U.S. Energy Information
Administration**](https://www.eia.gov/){target="_blank"}

-   Excellent information on electricity usage by location
    or region.

[**Data.Gov - Data and Statistics about the
U.S.**](https://data.gov/){target="_blank"}

-   Includes MANY topics including BLS, DOE, and EIA above

[**Yahoo
Finance**](https://finance.yahoo.com/){target="_blank"}

-   Stock data can be directly imported into R as an xts
    dataset

# Dashboard Examples

-   [**HW 5 - Part 1 - Quarto
    Publish**](https://penelope2040.quarto.pub/bua-455---hw-5/){target="_blank"}

-   [**HW 5 - Part 1 -
    RPubs**](https://rpubs.com/PeneLope_PE/1229274){target="_blank"}

<br>

-   [**Dashboard Example
    1**](https://rpubs.com/PeneLope_PE/1276088){target="_blank"}

    -   This group created a highly customized color scheme
        which is not required.

    -   Groups can use one of the many packaged dashboard
        themes that work really well.

    -   Overall, this is excellent work but dashboards
        should be less wordy.

<br>

-   [**Dashboard Example
    2**](https://rpubs.com/PeneLope_PE/1276097){target="_blank"}

    -   This group used the default theme which is fine, but
        it is easy to change.

    -   Overall, this is also excellent work, but again, a
        little too wordy.

    -   The perceptive depth of the analysis in the plots
        made this project stand out.

<br>

-   [**Dashboard Example
    3**](https://rpubs.com/PeneLope_PE/1276115){target="_blank"}

    -   This group used a packaged dashboard theme and
        complementary color pallettes that enhanced their
        dashboard.

    -   This group also did an excellent job of editing the
        text and almost all of the pages are well arranged.

    -   The data management in the dashboard was excellent.
        If this was a graduate class, I would encourage the
        students to go further with the analyses. I also
        might suggest more advanced plots and subsequent
        data exploration suggested by the work they had
        already completed.

<br>

-   [**Dashboard Example
    4**](https://rpubs.com/PeneLope_PE/1300146){target="_blank"}

    -   This group did some interesting work and I am
        including this example because it corresponds to the
        example memos provided later in the semester.

<br>

-   [**Presentation to 6th Grade
    Class**](https://peneloopy.github.io/sixth_grade_2025/survey_dashboard.html#){target="_blank"}

    -   In May of 2025, I asked two sixth grade classes to
        complete an anonymous survey and then I created a
        dashboard from their responses using the same
        criteria required for your projects.

    -   This dashboard is posted on GitHub.

    -   I will be using GitHub as the sole platform in a
        course I am developing, **Interactive Exploratory
        Data Analysis**.
